---
title: 'Blogdown 101: Enabling emoticons'
author: Nathan Brouwer
date: '2021-07-01'
slug: blogdown-101-emoticons
categories:
  - blogdown
tags:
  - blogdown
  - R
math: no
meta: yes
toc: yes
editor_options: 
  chunk_output_type: console
---


<!--more-->

```{r, echo = F}
library(blogdown)
library(tufte)
```



To enable emoticons add this to the header of config.toml

```{r, eval = F}
enableEmoji = true
```

For an example,  Maelle Salmon's config.toml on [GitHub](https://github.com/maelle/simplymaelle/blob/main/config.toml)
 
I've never used emoticons in RMarkdown and wasn't able initially to get them to show up usuing syntax like this:
:exclamation:

Using the [emo](https://github.com/hadley/emo) package worked.   As of Dec. 2020 emo isn't on CRAN but is available from GitHub
```{r, eval = F}
devtools::install_github("hadley/emo")
emo::ji("poop")
```


For some more info on blogdown and emoticons see [here](https://bookdown.org/yihui/blogdown/configuration.html).

A useful cheatsheet is [here](https://www.webfx.com/tools/emoji-cheat-sheet/).

To call an emoji in-line use this:

```{r, eval = F}
`r emo::ji("sos")` 
```

And you'll get this:
`r emo::ji("sos")` 


I'd like to figure out how to render them in margin notes but haven't figure it out yet.

```{r fig.margin=TRUE}
plot(cars)
```


```{r, echo = F, eval = F}
#I'm thinking about using emjoi in Tufte side-notes to signal certain themes:

# Protip
emo::ji("mortar_board")

# plotting info
emo::ji("chart_with_upwards_trend")
emo::ji("bar_chart")

# alternate code
emo::ji("floppy_disk")

# math note
emo::ji("heavy_division_sign")
emo::ji("heavy_plus_sign")
emo::ji("heavy_multiplication_x")

# practice
emo::ji("repeat")

# seeding an idea I'll be coming back to
emo::ji("seedling")


# package
emo::ji("package")

# warning
emo::ji("warning")

# help!
emo::ji("sos")

# details
emo::ji("mag")

# connection to other ideas
emo::ji("link")
```


Other emojis

```{r}
emo::ji("exclamation")
emo::ji("ballot_box_with_check")
emo::ji("no_entry_sign")
emo::ji("memo")
emo::ji("open_file_folder")
emo::ji("mega")
emo::ji("computer")
emo::ji("muscle")
emo::ji("question")
emo::ji("twisted_rightwards_arrows")
emo::ji("hourglass")
```






