---
title: Key files for the tuftesque hugo theme
author: Nathan Brouwer
date: '2020-12-04'
slug: key-files-for-the-tuftesque-hugo-theme
categories:
  - blogdown
tags:
  - Hugo
  - tuftesque
  - blogdown
math: no
meta: yes
toc: yes
draft: no
excerpt: "Key files for users to change to personal as tuftesque-themed blogdown site"
editor_options: 
  chunk_output_type: console
---


<!--more-->

lobrowR uses the (tuftesque)[https://github.com/nstrayer/tuftesque] Hugo theme.  This post outline key files I changed to set up the blog.  For more information see the [blog post](https://livefreeordichotomize.com/2017/07/13/introducing-the-tuftesque-blogdown-theme/) by Nick Stayer introducing the theme.

## The tuftesque theme

(lithium)[https://themes.gohugo.io/hugo-lithium-theme/] is the default Hugo theme for blogdown.  Its a nice, minimal theme useful for code-centered blogs.  The [hugo-tufte](https://github.com/shawnohare/hugo-tufte) theme implements [Edward Tufte's](https://en.wikipedia.org/wiki/Edward_Tufte) style and formatting as a Hugo theme, and has been further adapted as (tuftesque)[https://github.com/nstrayer/tuftesque] by Nick Stayer for use on the blog [Live Free or Dichotomize](https://livefreeordichotomize.com/), which he contributes to along with [Lucy D'Agostino McGowan](https://github.com/LucyMcGowan).  ([Dr. Lucy D'Agostino McGowan](https://www.lucymcgowan.com/) - also of the podcast [Casual Inference](https://casualinfer.libsyn.com/) fame - appears to be the primary contributor currently.)





## Changing the logo file

tuftesque defaults to a picture of Tin-tin.  The logo.png files shows up in several places, which is confusing to a new user.  

```{r,eval = F}
/themes/tuftesque/static
/public
```

Key to know is that /public shows up when you build the site for the first time using blogdown::build_site() and is where your rendered site lives. Changes to the underlying website structure therefore occur within theme/tuftesque and are implemented in flies in /public.

The default logo is 111 kb and 320×320.  Blogdown doesn't seem to re-size image on the fly and so I have only been able to change the image when a new images is approximately the same size.

To change the logo, add a new .png logo file to /themes/tuftesque/static,  then change what logo file is called up in the config.toml (which is in blogs root directory) under "[params.logo]"

```{r, eval = F}
    [params.logo]
    #url = "logo.png"
    url = "logo2.png"
```

I recommend commenting out the original call to the logo.png and leaving the original file so you can troubleshoot any errors or issues.

Run blogdown:::preview_site(startup = TRUE) to see if it worked.
```{r eval = F}
blogdown:::preview_site(startup = TRUE)
```

## Landing page banner

The banner at the top of the landing page lives in

/themes/tuftesque/static/images

Original tuftesque banner is 382 kb 1280×427.  I didn't experiment with anything else but set saved my file as exactly 1280×427.  (The code I used for my random-walk banner is at the end of the post).


## The authors.toml file

authors.toml adds a mini author biography at the end of each post.
Create authors.toml in themes/tuftesque/data/.  Below is my original file.

```{r, eval = F}
["Nathan Brouwer"]
  name =  "Nathan Brouwer"
  bio = "Nathan like data."
  location =  "Pittsburgh, PA"
  website =  "https://www.biology.pitt.edu/person/nathan-brouwer"
  thumbnail = "https://avatars2.githubusercontent.com/u/7841427?s=400&v=4"
  twitter = "https://twitter.com/lobrowr"
  github = "https://github.com/brouwern"
```

"thumbnail" is a picture of author.  The default is 29kb and 400×400;  I was able to use the link from my GitHub profile picture and didn't check the resolution of my file, and have any issues.

Note that the icon for GitHub that appears in the rendered author bio info doesn't seem to work currently.



## Change the background color

The default background is a creamy white.  I changed it to white.  Instructions are on the GitHub (landing page)[https://github.com/nstrayer/tuftesque] for the theme.  

In config.toml just 
backgroundcolor = "yourcolorchoice"
to the [params] field, eg.

```{r, eval = F}
[params]
    ...
    backgroundcolor = "white"
    ...
```


## Code to generate a random walk banner

Here's how I generate my banner.

```{r}
x <- 1:100
y <- rep(NA,100)
y[1] <- 100
lambda1 <- rnorm(n =length(y), mean = 1, sd = 0.2)
for(i in 2:length(y)){
  y[i] <- y[i-1]*lambda1[i]
}

par(mar = c(0,0,0,0))
plot(y ~ x, ylim = c(20,220),
     xlim = c(0,15),
     type = "l",yaxt="n", xlab="", ylab="",xaxt="n", 
     bty= "n", lwd = 5, lty = 1)

its <- 20
for(j in 1:its){
 y[1] <- 100

for(i in 2:length(y)){
  lambda <- 1
  lambda <- rnorm(n =1, mean = 1, sd = 0.1+i/1000+1/lambda/1000)
  y[i] <- y[i-1]*lambda
}

points(y ~ x, ylim = c(0,200), lwd = 2, type = "l", col = j) 
}

mtext(text = "  lobrowR",side = 3,line = -6, cex = 8,adj = 0)

```





