---
title: Key Files for the Tuftesque Hugo Theme
author: Nathan Brouwer
date: '2020-12-04'
slug: key-files-for-the-tufteque-hugo-theme
categories:
  - blogdown
tags:
  - Hugo
  - tuftesque
  - blogdown
math: no
meta: yes
toc: yes
exert: "Key files for users to change with the Tuftesque theme"
---


<!--more-->

Tufteque based on lithium - same key files?

## logo

Tufteque defaults to Tin-tin.
The logo.png files shows up in several places
It is 111 kb and 320×320

Blogdown doesn't seem to re-size image on the fly and so I have only been able to change the image when a new images is approximatley the same size.

```{r,eval = F}
/themes/tuftesque/static
/public
```



/public shows up when you build the site for the first time using blogdown::build_site().  This is where files are either copied to or built for actually posting the site.

To change the logo, add a new logo to /themes/tuftesque/static then change what logo file is called up in the config.toml file under "[params.logo]"

```{r, eval = F}
    [params.logo]
    #url = "logo.png"
    url = "logo2.png"
```


Run blogdown:::preview_site(startup = TRUE) to see if it worked.
```{r eval = F}
blogdown:::preview_site(startup = TRUE)
```






