---
title: Key files for the tuftesque hugo theme
author: Nathan Brouwer
date: '2020-12-04'
slug: key-files-for-the-tuftesque-hugo-theme
categories:
  - blogdown
tags:
  - Hugo
  - tuftesque
  - blogdown
math: no
meta: yes
toc: yes
draft: yes
excerpt: "Key files for users to change to personal as tuftesque-themed blogdown site"
editor_options: 
  chunk_output_type: console
---


<!--more-->

tuftesque based on lithium - same key files

## logo

tuftesque defaults to a picture of Tin-tin.
The logo.png files shows up in several places
It is 111 kb and 320×320

Blogdown doesn't seem to re-size image on the fly and so I have only been able to change the image when a new images is approximatley the same size.

```{r,eval = F}
/themes/tuftesque/static
/public
```



/public shows up when you build the site for the first time using blogdown::build_site().  This is where files are either copied to or built for actually posting the site.

To change the logo, add a new logo to /themes/tuftesque/static then change what logo file is called up in the config.toml file under "[params.logo]"

```{r, eval = F}
    [params.logo]
    #url = "logo.png"
    url = "logo2.png"
```


Run blogdown:::preview_site(startup = TRUE) to see if it worked.
```{r eval = F}
blogdown:::preview_site(startup = TRUE)
```

##  Banner

/themes/tuftesque/static/images

Original tufteque banner:
382 kb
1280×427

Cod to generate a random walk

```{r}
x <- 1:100
y <- rep(NA,100)
y[1] <- 100
lambda1 <- rnorm(n =length(y), mean = 1, sd = 0.2)
for(i in 2:length(y)){
  y[i] <- y[i-1]*lambda1[i]
}

par(mar = c(0,0,0,0))
plot(y ~ x, ylim = c(20,220),
     xlim = c(0,15),
     type = "l",yaxt="n", xlab="", ylab="",xaxt="n", 
     bty= "n", lwd = 5, lty = 1)

its <- 20
for(j in 1:its){
 y[1] <- 100

for(i in 2:length(y)){
  lambda <- 1
  lambda <- rnorm(n =1, mean = 1, sd = 0.1+i/1000+1/lambda/1000)
  y[i] <- y[i-1]*lambda
}

points(y ~ x, ylim = c(0,200), lwd = 2, type = "l", col = j) 
}

mtext(text = "  lobrowR",side = 3,line = -6, cex = 8,adj = 0)

```


## authors.toml

Add this to add a mini author biography at the end of each post.




