---
title: 'Example plot: means with error bars (error plot)'
author: Nathan Brouwer
date: '2021-02-01'
slug: ''
categories: []
tags: []
math: no
meta: no
toc: no
---


<!--more-->


```{r, echo = F}
# library(tufte)
# library(formatR)
# 
# library(readxl)
# 
# library(ggpubr)
# library(cowplot)
# 
# library(lme4)
# 
# library(dplyr)
# library(plotrix)
```

```{r  echo = F}
# file. <- "Meillere_et_al_2015_telomere_data.xlsx"
# 
# #load data
# parus_telos_nest <- readxl::read_xlsx(file.,sheet = "orig")
# 
# # clean up data
# parus_telos_nest$cort <- as.numeric(parus_telos_nest$cort)
# parus_telos_nest$Group <- with(parus_telos_nest,
#                        paste(treatment,sex))
# 
# parus_telos_nest$Group <- gsub("control","Control",parus_telos_nest$Group)
# parus_telos_nest$Group <- gsub("disturbed","Noise exposure",parus_telos_nest$Group)
# 
# parus_telos_nest$Group <- gsub(" F","\nFemales",parus_telos_nest$Group)
# parus_telos_nest$Group <- gsub(" M","\nMales",parus_telos_nest$Group)
# parus_telos_nest$cort.log <- log(parus_telos_nest$cort)
# parus_telos_nest$telomere.length.log <- log(parus_telos_nest$telomere.length)
# 
# 
# names(parus_telos_nest) <- gsub("treatment",
#                         "Treatment",
#                         names(parus_telos_nest))
# 
# names(parus_telos_nest) <- gsub("sex",
#                         "Sex",
#                         names(parus_telos_nest))
```

```{r}
# This script makes example of different kinds of plots using the telomere data 
# of Meillere et al 2015.

# Prep stuff
## libraries
library(ggpubr)
library(cowplot)
library(dplyr)
library(docxtractr)

library(plotrix) # for std.err


file.url <-  "https://royalsocietypublishing.org/action/downloadSupplement?doi=10.1098%2Frsbl.2015.0559&file=rsbl20150559supp1.docx"

### Works with the URL too
file.docx <- docxtractr::read_docx(file.url)
docx.table.list <- docxtractr::docx_extract_all_tbls(docx = file.docx)


## Call each table from list
### nestlings
parus_telos_nest <- docx.table.list[[1]]



# clean up data
parus_telos_nest$cort <- as.numeric(parus_telos_nest$cort)
parus_telos_nest$telomere.length <- as.numeric(parus_telos_nest$telomere.length)

parus_telos_nest$Group <- with(parus_telos_nest,
                       paste(treatment,sex))

parus_telos_nest$Group <- gsub("control","Control",parus_telos_nest$Group)
parus_telos_nest$Group <- gsub("disturbed","Noise exposure",parus_telos_nest$Group)

parus_telos_nest$Group <- gsub(" F","\nFemales",parus_telos_nest$Group)
parus_telos_nest$Group <- gsub(" M","\nMales",parus_telos_nest$Group)
parus_telos_nest$cort.log <- log(parus_telos_nest$cort)
parus_telos_nest$telomere.length.log <- log(parus_telos_nest$telomere.length)


names(parus_telos_nest) <- gsub("treatment",
                        "Treatment",
                        names(parus_telos_nest))

names(parus_telos_nest) <- gsub("sex",
                        "Sex",
                        names(parus_telos_nest))

## Re-code data
dat.focal.cols <- parus_telos_nest[,c("Treatment","Sex","telomere.length","cort")]
dat.foc.cols.rnd <- dat.focal.cols
dat.foc.cols.rnd$telomere.length <- dat.foc.cols.rnd$telomere.length %>% round(1)
dat.foc.cols.rnd <- dat.foc.cols.rnd  %>% dplyr::arrange(Treatment,telomere.length)
```



```{r, echo = F, message = F, warning = FALSE}

#parus_telos_nest$Sex <- ifelse(parus_telos_nest$Sex == "F","Female","Male")

# parus_telos_nest$Treatment <- factor(parus_telos_nest$Treatment, 
#                    levels = c("control","disturbed",""))

means <- parus_telos_nest %>%
 group_by(Treatment) %>%
  summarize(mean = mean(telomere.length),
            SD = sd(telomere.length),
            SE = std.error(telomere.length))


means$CI.top <- means$mean+2*means$SE
means$CI.bottom <- means$mean-2*means$SE


```



```{r  echo = F}
ggplot(data = means,
       aes(y = mean,
           x = Treatment,
           color = Treatment,
           shape = Treatment)) +
  geom_errorbar(aes(ymin = CI.top,
                    ymax = CI.bottom),
                width = 0,
                size = 2,
                color = "black") + 
    geom_point(size = 8) +
  scale_x_discrete(limits = c("","control","disturbed","")) +
  theme(legend.position="none") +
   ggtitle("Error plot: means & error bars\n") +
   xlab("Predictor: Experimental group\n") +
   ylab("Response: Telomere length (kb)\n") +
  
  #Full CI
   geom_segment(x = 2.15, 
                xend = 2.05,
                y = means[1,"CI.top"][[1]],
                yend = means[1,"CI.top"][[1]],
               color = "black",
               arrow = arrow(length = unit(0.03, "npc"))) +
   geom_segment(x = 2.15, 
                xend = 2.05,
                y = means[1,"CI.bottom"][[1]],
                yend = means[1,"CI.bottom"][[1]],
               color = "black",
               arrow = arrow(length = unit(0.03, "npc"))) +
  geom_segment(x = 2.15, 
                xend = 2.15,
                y = means[1,"CI.top"][[1]],
                yend = means[1,"CI.bottom"][[1]],
               color = "black") +
  annotate(geom = "text",
           x = 2.5,
           y = means[1,"mean"][[1]],
           label = "Full confidence\ninterval\n(+/- 2*SE)") +
  # Upper CI
   geom_segment(x = 1.75, 
                xend = 1.95,
                y = means[1,"CI.top"][[1]],
                yend = means[1,"CI.top"][[1]],
               color = "black",
               arrow = arrow(length = unit(0.03, "npc"))) +
  annotate(geom = "text",
           x = 1.25,
           y = means[1,"CI.top"][[1]],
           label = "Top of CI\n(Y = mean+2*SE)")+
  
  #Lower CI
   geom_segment(x = 1.75, 
                xend = 1.95,
                y = means[1,"CI.bottom"][[1]],
                yend = means[1,"CI.bottom"][[1]],
               color = "black",
               arrow = arrow(length = unit(0.03, "npc"))) +
  annotate(geom = "text",
           x = 1.25,
           y = means[1,"CI.bottom"][[1]],
           label = "Bottom of CI\n(Y = mean - 2*SE)") +
  
  # Mean
  geom_segment(x = 1.75, 
                xend = 1.9,
                y = means[1,"mean"][[1]],
                yend = means[1,"mean"][[1]],
               color = "black",
               arrow = arrow(length = unit(0.03, "npc")))  +
  annotate(geom = "text",
           x = 1.5,
           y = means[1,"mean"][[1]],
           label = "Mean") +
  #Half CI
   geom_segment(x = 3.25, 
                xend = 3.15,
                y = means[2,"CI.top"][[1]],
                yend = means[2,"CI.top"][[1]],
               color = "black",
               arrow = arrow(length = unit(0.03, "npc"))) +
   geom_segment(x = 3.25, 
                xend = 3.15,
                y = means[2,"mean"][[1]],
                yend = means[2,"mean"][[1]],
               color = "black",
               arrow = arrow(length = unit(0.03, "npc"))) +
  geom_segment(x = 3.25, 
                xend = 3.25,
                y = means[2,"CI.top"][[1]],
                yend = means[2,"mean"][[1]],
               color = "black") +
  annotate(geom = "text",
           x = 3.75,
           y = 0.98*means[2,"CI.top"][[1]],
           label = "Upper error bar of\nconfidence interval\n(mean to \nmean + 2*SE)")  +
  
  #text
  annotate(geom = "text",
           x = 3,
           y = 1.25,
           label = "An error bar extends from the
mean up or down to a certain
value.

A confidence interval refers to
the whole span from the top
of the top bar to the bottom
of the lower bar.",hjust =0)  +
  ylim(0.95,1.35) +
  theme_bw() + theme(legend.position = "none")
```

An **error bar** represents variation around a mean value or uncertainty we  have in our estimate of the mean.  Variation is represented by the **standard deviation (SD)** and characterizes just variation in the raw data.  Uncertainty in the estimate is represented by the **standard error (SE)** or a **95% confidence interval (95% CI)**.

When we refer to "an error bar" we usually are referring to a single bar extending either above or below the mean.  When we refer to a confidence interval we are referring to both bars.  A confidence interval is the full span from the top of the upper error bar to the bottom of the lower bar.  There is therefore a subtle difference between referring to error bars from standard deviations or standard error and the 95% confidence.

One thing to remember is that the bars shown in an error plot or barplot are completely different than those in a **boxplot**.  Its useful to remember that the full name of a boxplot is a  "box and whisker plot".  The bars extending above the boxes in a boxplot are *not* error bars.  The top of the upper "**whisker**" goes up to the maximum value of the data (except when there are outliers); the bottom of the lower whisker goes down to the minimum value (except when there are outliers).  These whiskers span upper and lower quartiles and are meant to capture aspects of the **distribution** of the data.

Data are from 
Meillere et al 2015. Traffic noise exposure affects telomere length in nestling house sparrows.  Proceedings of the Royal Society B.  11: 20150559 https://doi.org/10.1098/rsbl.2015.0559
