---
title: Building a basic blogdown site with the tuftesque theme
author: Nathan Brouwer
date: '2020-12-17'
slug: building-a-basic-blogdown-site-workflow
categories:
  - blogdown
tags:
  - Hugo
  - blogdown
  - tuftesque
math: no
meta: no
toc: no
exert: "Workflow for setting up blogdown with tuftesque"
---


<!--more-->

Had some problems, hopefully this workflow will work.

## Uninstall Hugo

I was having some issues so I re-installed RStudio and Hugo.  


https://github.com/rstudio/blogdown/issues/324
Command line:
brew remove hugo

Where hugo hides:
~/Library/Application\ Support/Hugo/
/usr/local/Cellar/hugo
/usr/local/bin/hugo 


## Create site

blogdown::new_site(
  dir = "~/Box Sync/R/lobrowr", # need to set directory to be made
  install_hugo = TRUE,
  format = "toml",
  sample = TRUE,
  theme = "yihui/hugo-lithium", # use original theme
  hostname = "github.com",
  theme_example = TRUE,
  empty_dirs = FALSE,
  to_yaml = TRUE,
  serve = interactive()
)

## Set directory with site as package

usethis::create_project(
  path = "~/Box Sync/R/lobrowr",
  rstudio = rstudioapi::isAvailable(),
  open = rlang::is_interactive()
)

## stop sever to prevent hiccups ...

blogdown::stop_server()

## install the them

blogdown::install_theme('nstrayer/tuftesque',force = TRUE)

## preview again - new theme now implemented 

blogdown:::preview_site(startup = TRUE)



## Set up Hugo page bundles

Followed advice here
 https://alison.rbind.io/post/2019-02-21-hugo-page-bundles/
 
Edited R profile
usethis::edit_r_profile(scope = "project")


## Associate with git

https://happygitwithr.com/existing-github-last.html

usethis::use_git()

## Link with GitHub

git remote add origin https://github.com/brouwern/lobrowr.git

git push --set-upstream origin master


## Change the logo 

See the post "Key files for the tufteque hugo theme" for info on this

## Change the background color

From the GitHub landing page for the theme
https://github.com/nstrayer/tuftesque

in config.toml add 
backgroundcolor = "yourcolorchoice"
to the [params]

```{r, eval = F}
[params]
    ...
    backgroundcolor = "white"
    ...
```


## Author info

As per

Add file authors.toml to themes/tufteque/data/

```{r, eval = F}
["Nathan Brouwer"]
  name =  "Nathan Brouwer"
  bio = "Nathan like data."
  location =  "Pittsburgh, PA"
  website =  "https://www.biology.pitt.edu/person/nathan-brouwer"
  thumbnail = "https://avatars2.githubusercontent.com/u/7841427?s=400&v=4"
  twitter = "https://twitter.com/lobrowr"
  github = "https://github.com/brouwern"
```

thumbnail is a picture of author, 29kb and 400×400.  I was able to use the link from my GitHub profile picture.

The icon for GitHub doesn't seem to work currently.

## Other potentially useful commands

blogdown::hugo_version()
blogdown::update_hugo()

blogdown::install_hugo(version="latest", force=TRUE)


