---
title: Building a basic blogdown site with the tuftesque theme
author: Nathan Brouwer
date: '2020-12-17'
slug: building-a-basic-blogdown-site-workflow
categories:
  - blogdown
tags:
  - Hugo
  - blogdown
  - tuftesque
math: no
meta: no
toc: no
draft: no
excerpt: "Workflow for setting up blogdown with tuftesque"
---


<!--more-->


I initially set this blog up via the GUI interface and instructions on Nick Strayer's blog post the (tuftesque Hugo theme)[https://livefreeordichotomize.com/2017/07/13/introducing-the-tuftesque-blogdown-theme/] and Dan Quintana's on the [Academic Hugo theme ](https://www.dsquintana.blog/free-website-in-r-easy/).  I broke my initial set up and had to troubleshoot some things. The workflow below ended up working


## Uninstall Hugo

As always, when things stop working its good to uninstall and re-install things.  I doubt this was contributing to the problem.  Uninstalling Hugo took a few Google searches.  From the command line according to (this)[https://github.com/rstudio/blogdown/issues/324]

```{r, eval = F}
brew remove hugo
```

Hugo can show up a few places so I double checked these locations as per the post linked above:
```{r, eval = F}
~/Library/Application\ Support/Hugo/
/usr/local/Cellar/hugo
/usr/local/bin/hugo 
```

## Install Hugo

I then re-installed Hugo

```{r, eval = F}
blogdown::install_hugo(version="latest", force=TRUE)
```




## Create site

Re-install stuff didn't help, so I started rebuilding the site.  I then had problems with the RStudio's GUI interface for creating a blogdown site, so I did it with the code below.  Note that I started with the default lithium theme, which I then install tuftesque over (below).
```{r, eval = F}
blogdown::new_site(
  dir = "~/Box Sync/R/lobrowr", # need to set directory to be made
  install_hugo = TRUE,
  format = "toml",
  sample = TRUE,
  theme = "yihui/hugo-lithium", # use original theme
  hostname = "github.com",
  theme_example = TRUE,
  empty_dirs = FALSE,
  to_yaml = TRUE,
  serve = interactive()
)
```


## Set directory with site as package

I then defined the folder containing the blog as an R project.  (This is done automatically when the RStudio GUI is used, but again, that was giving me issue).

```{r, eval = F}
usethis::create_project(
  path = "~/Box Sync/R/lobrowr",
  rstudio = rstudioapi::isAvailable(),
  open = rlang::is_interactive()
)

```

## Stop serving the preview site to prevent hiccups 

I'm currently working on a Mac.  Previously -- and on a PC -- I had lots of issues with the preview site generated by RStudio.  I'm having no such problems now, but stopping the sever can't hurt.

```{r, eval = F}
blogdown::stop_server()
```




## Install the tuftesque theme

A new theme can be install using install_theme().  One of the issues I had with using the RStudio GUI to generate the new blog was that I was getting an error related to having installed the tuftesque theme before.  force = TRUE re-installs it over top of an existing installation. (Not sure why there was a problem with this i the first place).

```{r, eval = F}
blogdown::install_theme('nstrayer/tuftesque',
                        force = TRUE)
```


## Preview again - hurray -  new theme now implemented!

I don't know what's going on under the hood, but installing tuftesque causes it to be used as the theme for the blog rather than the default lithium.  THere's no need to declare which theme to use.

```{r, eval = F}
blogdown:::preview_site(startup = TRUE)
```



## Set up Hugo page bundles

**Hugo page bundles** make keeping things organized easy.  I followed the advice from Allison Hill from (here)[https://alison.rbind.io/post/2019-02-21-hugo-page-bundles/] and edited the R profile.

A project-specific R profile can be created like this:
```{r, eval = F}
usethis::edit_r_profile(scope = "project")
```



## Associate with git

The RStudio GUI for creating a new blog doesn't allow you to set up Git or associate with GitHub.  I followed Jenny Bryan's instructions from (here)[https://happygitwithr.com/existing-github-last.html] and did this by hand.  The command to set its just

```{r, eval = F}
usethis::use_git()
```


## Link with GitHub

To associate with GitHub I used the command line:

```{r, eval = F}
git remote add origin https://github.com/brouwern/lobrowr.git
git push --set-upstream origin master
```







