---
title: Scraping Restriction Enzyme Data from Wikipedia
author: Nathan Brouwer
date: '2021-02-17'
slug: scraping-restriction-enzyme-data-from-wikipedia
categories: 
  - R
tags:
  - bioinformatics
math: no
meta: yes
toc: yes
editor_options: 
  chunk_output_type: console
---


<!--more-->

Challenge: Get this table and extract biological information from it
(show table)

## Tasks

1. Scrape table from Wikipedia; very easy - poweR tool.
1. Scrape related tables from multiple Wiki pages; easy, once you know for() loops
1. (split recognition sequence)
1. (split cut and overhang)
1. (deal with Isoschizomers)
1. Extract length of recognition sequence
1. Extract Neoschizomers (same recog, different cut - this is shown with green font - doh!)

what does this mean: "Isoschizomers indicated in white font and grey background correspond to enzymes not listed in the current lists, as in this not listed enzyme" (example..)

https://rlang.io/scraping-tables-from-wikipedia-for-visualizing-climate-data/


Packages 

```{r, eval = T}
# install.packages("rvest")
#install.packages("data.table")
# install.packages("magrittr")
# install.packages("ggthemes")

library("rvest") # Scraping
library("data.table") # Melting
library("dplyr") # Manipulating
library("ggplot2") # Plotting
```

Assign the URL for the Wiki page to an object
```{r}
url <- "https://en.wikipedia.org/wiki/List_of_restriction_enzyme_cutting_sites:_A#Whole_list_navigation"
```

Read the URL with read_html() and extract the table with  html_nodes(..., "table")

```{r}
wiki <- read_html(url)
tables <- html_nodes(wiki, "table")
```

This produces an "xml_nodeset" object containing each table
```{r}
class(tables)
```

There happen to be 3 tables in on this Wikipedia page
```{r}
length(tables)
```


Each table is indicated within the object by a rather cryptic string 
```{r}
tables
```

Each table is accessed via bracket notation
```{r}
tables[1]
```

and is extracted using html_table().  fill = TRUE fills in NAs as needed.
```{r}
table1.list <- html_table(tables[1], fill = TRUE)
```

The output of html_table() is a list
```{r}
class(table1.list)
```

To extract the dataframe within the list use double-bracket notation to call up the first (and in this case only) item of the list:

```{r}
table1 <- table1.list[[1]]
```

This can be done in one step by appending the [[1]] just outside the call to html_table()

```{r}
table2 <- rvest::html_table(tables[2], fill = TRUE)[[1]]
table3 <- rvest::html_table(tables[3], fill = TRUE)[[1]]
```


```{r}
table2 <- data.frame(table2)

table3 <- data.frame(table3)

dim(table2)
dim(table3)

```



A Ba-Bc Bd-Bp Bsa-Bso Bsp Bsr-Bss Bst Bsu-Bv –CDEFGHIKLMNOPRSTUVXYZ– External links See also

```{r}

```

